Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-28 cav2017/cav2017.tex
edits

The models we consider have two sources of error due to numeric
representation.  The first is the numeric error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$ and $x_k$.  The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
fixed-point representation of numbers, and briefly describe the errors
introduced.  A formal discussion is in
Appendix~\ref{appendix:numerical_errors}.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesizer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
The input specification provided to the program synthesizer is of the form
%$\exists P .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ 
$\exists P .  \forall x.  \sigma(x, P)$, 
where $P$ ranges over functions, 
$x$ ranges over ground terms, 
and $\sigma$ is a quantifier-free formula.  
We interpret the ground terms over some finite domain~$\mathcal{D}$.
%
The design of our synthesizer consists of two phases, an inductive
synthesis phase and a validation phase, which interact via a finite
set of test vectors {\sc inputs} that is updated incrementally.  Given
the aforementioned specification $\sigma$, the inductive synthesis
procedure tries to find an existential witness $P$ satisfying
the specification $\sigma(x, P)$ for all $x$ in {\sc
  inputs} (as opposed to all $x \in \mathcal{D}$).
%
If the synthesis phase succeeds in finding a witness~$P$, this
witness is a candidate solution to the full synthesis formula.  We
pass this candidate solution to the validation phase, which checks
whether it is a full solution (i.e., $P$ satisfies the
specification $\sigma(x, P)$ for all
$x\in\mathcal{D}$).  If this is the case, then the algorithm
terminates.  Otherwise, additional information is provided to the
inductive synthesis phase in the form of a new counterexample that is
added to the {\sc inputs} set and the loop iterates again.
More details about the general architecture of the synthesizer can
be found in~\cite{DBLP:conf/lpar/DavidKL15}.
%% It is worth noting that each iteration of the loop adds a new input to
%% the finite set $\text{\sc inputs}$ that is used for synthesis.  Given
%% that the full set of inputs $\mathcal{D}$ is finite, this means that
%% the refinement loop can only iterate a finite number of times.

@@

A stable control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this in general not the case when dealing with complex
eigenvalues.  Consider the closed-loop matrix, which updates the states
every discrete time step, and select the eigenvalue $\vartheta$ with
smallest (non-trivial) imaginary value.  At every consecutive time steps
$kT_s, (k+1)T_s$, the dynamics projected on the corresponding eigenspace
rotate of $\vartheta T_s$ radiants.  So taking $k{\leq}\overline{k}$ as the
ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k{\leq}\overline{k}$ steps
we have completed a full rotation, which results in a point closer to the
origin, and it is the completeness threshold.

@@

We have  used $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.
The following approximation errors will arise in mathematical operations and representation:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed-point domain as above. 
Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T$ where the error $ \delta_T=x\ \%_{c_m}\ \tilde x$, and $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
%
\item {\bf Rounding:} The following errors appear in basic operations.  Let
$c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be
the truncation errors caused by representing $c_1$ and $c_2$ in the
fixed-point domain as above.
%
\begin{enumerate}
%
\item Addition/Subtraction: these operations only propagate errors coming
from truncation of the operands, namely $\mathcal{F}_{\langle I,F
\rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$
with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
%
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) \cdot
\mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3|
\leq |\delta_{T1}\cdot\mathcal{F}_{\langle I,F \rangle}(c_2)|\allowbreak +
|\delta_{T2}\cdot\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where
$c_m=2^{-F}$ as above.
%
\item Division: the operations performed by our controllers in the FWL
domain do not include division.  However, we do use division in computations
at the precision of the plant.  Here the error depends on whether the
divisor is greater or smaller than the dividend.  $\mathcal{F}_{\langle I,F
\rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_{T3}$
where $\delta_{T3}$ is $(\delta_{T2}\cdot c_1 - \delta_{T1}\cdot
c_2)/(\delta_{T2}^2 - \delta_{T2} c_2)$,
%
%$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  (c_1 - \delta_{T1})/ (c_2 + \delta_{T2}) + c_m $, where $c_m=2^{-F}$ as above.
\end{enumerate}

>>>>

The models we consider have two sources of error due to numeric
representation.  The first is the numeric error introduced by the
fixed-point numbers employed to model the plant, i.e., to represent the
plant dynamics $A_d$, $B_d$ and $x_k$.  The second is the quantization error
introduced by the digital controller, which performs operations on
fixed-point numbers.  In this section we outline the notation for the
fixed-point representation of numbers, and briefly describe the errors
introduced.  A~formal discussion is in
Appendix~\ref{appendix:numerical_errors}.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{General architecture of the program synthesizer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\label{synthesizer-general}
The input specification provided to the program synthesizer is of the form
%$\exists P .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ 
$\exists P .\,  \forall x.\,  \sigma(x, P)$, 
where $P$ ranges over functions, 
$x$ ranges over ground terms, 
and $\sigma$ is a quantifier-free formula.  
We interpret the ground terms over some finite domain~$\mathcal{D}$.
%
The design of our synthesizer consists of two phases, an inductive
synthesis phase and a validation phase, which interact via a finite
set of test vectors {\sc inputs} that is updated incrementally.  Given
the aforementioned specification $\sigma$, the inductive synthesis
procedure tries to find an existential witness $P$ satisfying
the specification $\sigma(x, P)$ for all $x$ in {\sc
  inputs} (as opposed to all $x \in \mathcal{D}$).
%
If the synthesis phase succeeds in finding a witness~$P$, this
witness is a candidate solution to the full synthesis formula.  We
pass this candidate solution to the validation phase, which checks
whether it is a full solution (i.e., $P$ satisfies the
specification $\sigma(x, P)$ for all
$x\in\mathcal{D}$).  If this is the case, then the algorithm
terminates.  Otherwise, additional information is provided to the
inductive synthesis phase in the form of a new counterexample that is
added to the {\sc inputs} set and the loop iterates again.
More details about the general architecture of the synthesizer can
be found in~\cite{DBLP:conf/lpar/DavidKL15}.
%% It is worth noting that each iteration of the loop adds a new input to
%% the finite set $\text{\sc inputs}$ that is used for synthesis.  Given
%% that the full set of inputs $\mathcal{D}$ is finite, this means that
%% the refinement loop can only iterate a finite number of times.

@@

A stable control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this in general not the case when dealing with complex
eigenvalues.  Consider the closed-loop matrix, which updates the states
every discrete time step, and select the eigenvalue $\vartheta$ with
smallest (non-trivial) imaginary value.  At every consecutive time steps
$kT_s, (k+1)T_s$, the dynamics projected on the corresponding eigenspace
rotate of $\vartheta T_s$ radiants.  Thus, taking $k{\leq}\overline{k}$ as the
ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k{\leq}\overline{k}$ steps
we have completed a full rotation, which results in a point closer to the
origin, and it is the completeness threshold.

@@

We have  used $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number
$x$ represented in a fixed point domain, with $I$ bits representing the
integer part and $F$ bits representing the decimal part.  The smallest
number that can be represented in this domain is $c_m=2^{-F}$.  The
following approximation errors will arise in mathematical operations and
representation:
%
\begin{enumerate}

\item {\bf Truncation:} Let $x$ be a real number, and $\mathcal{F}_{\langle
I,F \rangle}(x)$ be the same number represented in a fixed-point domain as
above.  Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T$ where the
error $ \delta_T=x\ \%_{c_m}\ \tilde x$, and $\%_{c_m}$ is the modulus
operation performed on the last bit.

Thus, $\delta_T$ is the truncation error and it will propagate across
operations.
%
\item {\bf Rounding:} The following errors appear in basic operations.  Let
$c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be
the truncation errors caused by representing $c_1$ and $c_2$ in the
fixed-point domain as above.
%
\begin{enumerate}
%
\item Addition/Subtraction: these operations only propagate errors coming
from truncation of the operands, namely $\mathcal{F}_{\langle I,F
\rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3$
with $|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.
%
\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) \cdot
\mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3$ with $|\delta_3|
\leq |\delta_{T1}\cdot\mathcal{F}_{\langle I,F \rangle}(c_2)|\allowbreak +
|\delta_{T2}\cdot\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where
$c_m=2^{-F}$ as above.
%
\item Division: the operations performed by our controllers in the FWL
domain do not include division.  However, we do use division in computations
at the precision of the plant.  Here the error depends on whether the
divisor is greater or smaller than the dividend.  $\mathcal{F}_{\langle I,F
\rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_{T3}$
where $\delta_{T3}$ is $(\delta_{T2}\cdot c_1 - \delta_{T1}\cdot
c_2)/(\delta_{T2}^2 - \delta_{T2} c_2)$,
%
%$\mathcal{F}_{\langle I,F \rangle}(c_1) / \mathcal{F}_{\langle I,F \rangle}(c_2) =  (c_1 - \delta_{T1})/ (c_2 + \delta_{T2}) + c_m $, where $c_m=2^{-F}$ as above.
\end{enumerate}
